\chapter{Introduction}

Over approximately 18 years, this project evolved to focus on a number of
related topics, all tied to the nuclear analysis of fusion energy systems.

For the earliest years, the \gls{UW}'s effort was in support of the \gls{APEX}
study to investigate high power density first wall and blanket systems.  A
variety of design concepts were studied before this study gave way to a
design effort for a US \gls{TBM} to be installed in ITER.  Simultaneous to
this \gls{TBM} project, nuclear analysis supported the conceptual design of a
number of fusion nuclear science facilities that might fill a role in the path
to fusion energy.

Beginning in approximately 2005, this project added a component focused on the
development of novel radiation transport software capability in support of the
above nuclear analysis needs.  Specifically, a clear need was identified to
support neutron and photon transport on the complex geometries associated with
\gls{CAD}.  Following the initial development of the \gls{DAGMC} capability,
additional features were added, including unstructured mesh tallies and
multi-physics analysis such as the \gls{R2S} methodology for \gls{SDR}
prediction.

Throughout the project, there were also smaller tasks in support of the fusion
materials community and for the testing of changes to the nuclear data that is
fundamental to this kind of nuclear analysis.

\section{About this Report}

This report summarizes the major findings over the duration of the project.
For many sections, text has be excerpted from publications that described this
work in more detail, with references provided for those publications.  The
following chapter describes a variety of nuclear analysis efforts, followed by
a chapter on the development of improved predictive capability, and then one
chapter each on nuclear data improvements and fusion materials support.

The references are divided into two sections: those that serve as research
products of this project, in whole or in part, and those that are outside this
project, but related in some way.
